---
title: "CDC data exercise"
---

This data set is provided by the Division of Adolescent School Health(DASH) and can be found here: <https://data.cdc.gov/Youth-Risk-Behaviors/DASH-YRBSS-HS-Cigar-Use-by-total-grade-sex-race-se/bedg-mmpy/about_data>.

The Youth Risk Behavior Surveillance System (YRBSS) monitors six categories of priority health behaviors among youth and young adults: 1) behaviors that contribute to unintentional injuries and violence; 2) tobacco use; 3) alcohol and other drug use; 4) sexual behaviors related to unintended pregnancy and sexually transmitted infections (STIs), including human immunodeficiency virus (HIV) infection; 5) unhealthy dietary behaviors; and 6) physical inactivity. In addition, YRBSS monitors the prevalence of obesity and asthma and other priority health behaviors. This dataset contains national, state, and local data from 2015 that includes two aspects of sexual orientation – sexual identity and sex of sexual contacts. Additional information about the YRBSS can be found at [www.cdc.gov/yrbss](http://www.cdc.gov/yrbss).

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(here)
```

```{r,}
#Define data location using here() package
data_location = here::here("cdcdata-exercise","cdc.csv")

# Load Data
data = read_csv(data_location)
head(data)
str(data)

```

```{r}
cdc <- subset(data, select = c("YEAR","Subtopic", "ShortQuestionText", "Greater_Risk_Data_Value", "Lesser_Risk_Data_Value"))

str(cdc)
```

Summarizing Year

```{r}
# Frequency table
year_freq <- table(cdc$YEAR)

# Percentage table
year_percent <- prop.table(year_freq) * 100

# Print tables
print(year_freq)
print(year_percent)

# Bar plot
library(ggplot2)
ggplot(cdc, aes(x = factor(YEAR))) +
  geom_bar() +
  labs(title = "Frequency of YEAR", x = "YEAR", y = "Count")

```

Summarizing Subtopic

```{r}
# Frequency table
subtopic_freq <- table(cdc$Subtopic)

# Percentage table
subtopic_percent <- prop.table(subtopic_freq) * 100

# Print tables
print(subtopic_freq)
print(subtopic_percent)

# Bar plot
ggplot(cdc, aes(x = Subtopic)) +
  geom_bar() +
  labs(title = "Frequency of Subtopic", x = "Subtopic", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Summarizing ShortQuestionText

```{r}
# Frequency table
shortquestiontext_freq <- table(cdc$ShortQuestionText)

# Percentage table
shortquestiontext_percent <- prop.table(shortquestiontext_freq) * 100

# Print tables
print(shortquestiontext_freq)
print(shortquestiontext_percent)

# Bar plot
ggplot(cdc, aes(x = ShortQuestionText)) +
  geom_bar() +
  labs(title = "Frequency of ShortQuestionText", x = "ShortQuestionText", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Summarizing Greater Risk Data Value

```{r}
# Descriptive statistics
greater_risk_summary <- summary(cdc$Greater_Risk_Data_Value)
greater_risk_mean <- mean(cdc$Greater_Risk_Data_Value, na.rm = TRUE)
greater_risk_sd <- sd(cdc$Greater_Risk_Data_Value, na.rm = TRUE)

# Print statistics
print(greater_risk_summary)
cat("Mean:", greater_risk_mean, "\n")
cat("Standard Deviation:", greater_risk_sd, "\n")

# Histogram and density plot
ggplot(cdc, aes(x = Greater_Risk_Data_Value)) +
  geom_histogram(aes(y = ..density..), binwidth = 1, fill = "blue", alpha = 0.7) +
  geom_density(color = "red") +
  labs(title = "Distribution of Greater_Risk_Data_Value", x = "Greater_Risk_Data_Value", y = "Density")
```

Summarizing Lesser Risk Data Value

```{r}
# Descriptive statistics
lesser_risk_summary <- summary(cdc$Lesser_Risk_Data_Value)
lesser_risk_mean <- mean(cdc$Lesser_Risk_Data_Value, na.rm = TRUE)
lesser_risk_sd <- sd(cdc$Lesser_Risk_Data_Value, na.rm = TRUE)

# Print statistics
print(lesser_risk_summary)
cat("Mean:", lesser_risk_mean, "\n")
cat("Standard Deviation:", lesser_risk_sd, "\n")

# Histogram and density plot
ggplot(cdc, aes(x = Lesser_Risk_Data_Value)) +
  geom_histogram(aes(y = ..density..), binwidth = 1, fill = "blue", alpha = 0.7) +
  geom_density(color = "red") +
  labs(title = "Distribution of Lesser_Risk_Data_Value", x = "Lesser_Risk_Data_Value", y = "Density")
```
